Graph Generators: State of the Art and Open Challenges

### 2019-07-23 generators_SocialNetworks.tex
Consolidation of Sections 3-5

On-line social networks, like Facebook, Twitter, or LinkedIn, have become a
phenomenon used by billions of people every day and thus providing extremely
useful information for various domains. However, an analysis of such type of
graph has to cope with two problems: (1) availability of the data and (2)
privacy of the data. Hence, data generators which provide realistic synthetic
social network graphs are in a great demand.

In general, any analysis of social networks identifies their various specific
features~\cite{Chakrabarti:2006:GML:1132952.1132954}. For example, a social
graph often has high \emph{clustering coefficient}, i.e. the degree of
transitivity of a graph. Or, its diameter, i.e. the longest shortest path
amongst some fraction (e.g. 90\%) of all connected nodes, is usually low due to
weak ties joining faraway cliques.

Another important aspect of social networks is the community effect. A detailed
study of structure of communities in 70 real-world networks is provided, e.g.,
in ~\cite{Leskovec:2008:SPC:1367497.1367591}.
~\cite{Prat-Perez:2014:CSS:2621934.2621942} analyzed the structure of
communities (clustering coefficient, triangle participation ratio, bridges,
diameter, conductance and size) in both real-world graphs and outputs of
existing graph generators such as LFR~\cite{PhysRevE.78.046110} and the
LDBC-SNB~\cite{Erling:2015:LSN:2723372.2742786}. They found out that discovered
communities  in different graphs have common distributions and that communities
in a single graph have diverse nature and are difficult to fit with a single
model.

@@

\paragraph{Realistic Social Network}
~\cite{Barrett:2009:GAL:1995456.1995598} focused on the construction of
realistic social networks. For this purpose the authors combine both public and
private data sets with large-scale agent-based techniques. The process works as
follows: In the first step it generates a synthetic data by combining public and
commercial databases. In the second step, it determines a set of activity
templates. A 24-hour activity sequence including geolocations is assigned to
each synthetic individual. To demonstrate the approach, the authors create a
synthetic US population consisting of people and households together with
respective geolocations. For this purpose the authors combine simulation and
data fusion techniques utilizing various real-world data sources such as U.S.
Census data, responses to an activity survey or a time-use survey.
%Demographic information for each person and location, a minute-by-minute
%schedule of each person's activities, and the locations where these activities
%take place is generated by a combination of simulation and data fusion
%techniques.
The result is captured by a dynamic social contact network. Similar methods for
agent-based strategies have been reported
in~\cite{Bernstein:2013:SAS:2499604.2499609}.

@@

\paragraph{Cloning of Social Networks} ~\cite{Sukthankar-SocialInfo2014}
introduces two synthetic generators to reproduce two characteristics typically
observed in social networks: node features and multiple link types. Both
generators etend on the generator proposed by~\cite{wang2011leveraging}.
which starts with a small number of
nodes, and new nodes are added until the network reaches the required number. It
has two basic parameters: homophily and link density. A high \emph{homophily}
value reflects that links have higher chances to be established among the nodes with the
same label; these labels can be considered as being equivalent to community
membership.

The first proposed generator is Attribute Synthetic Generator (ASG), used for
reproducing the node feature distribution of standard networks and rewiring the
network to preferentially connect nodes that exhibit a high feature similarity.
The network is initialized with a group of three nodes, and new nodes and links
are added to the network based on link density, homophily, and feature
similarity. As new nodes are created, their labels are assigned based on the
prior label distribution. After the network has reached the same number of nodes
as the original social media dataset, each node initially receives a random
attribute assignment. Then a stochastic optimization process is used to move the
initial assignments closer to the target distribution extracted from social
media dataset using the Particle Swarm Optimization algorithm. The tuned
attributes are then used to add additional links to the network based on the
feature similarity parameter -- a source node is selected randomly and connected
to the most similar node. The second proposed generator is Multi-Link Generator
(MLG) further  uses link co-occurrence statistics from the original dataset to
create a multiplex network. MLG uses the same network growth process as ASG.
Based on the link density parameter, either a new node is generated with a label
based on the label distribution of the target dataset or a new link is created
between two existing nodes.

@@

\paragraph{Towards More Realistic Data} \cite{Nettleton2016} argued that the majority of existing works focuses on
topology generation which approximates the characteristics of a real social
network (such as a small graph diameter, skew degree
distribution, small average path length, and community structures), however,  this is usually done without any data. Hence, they
introduced a general stochastic modeling approach that enables the users to
populate a graph topology with data. The approach has three steps: (1) topology
generation (using R-MAT) plus community identification using the Louvain
method~\cite{1742-5468-2008-10-P10008} or usage of a real-world topology from
SNAP\footnote{\url{https://snap.stanford.edu/data/}}, (2) data definition
that describes distribution profiles, attribute value definitions, using a
parameterizable set of data propagation rules and affinities, and (3) data
population.

\paragraph{Strengths and Weaknesses of Social Network Generators}
Compared to more general graph generators, social network generators focus
mainly on reproducing intra- and inter-node feature correlations.
Among existing genrators, LDBC-SNB and S3G2 look
like the most advanced ones in terms of the complexity of the generated graph
and the amount of features and correlations they can generate, while providing a
large degree of scalability. Their generation process is based on input dictionaries and have configuration files that
allow tweaking parameters many parameters of the generated graphs, but their schema is mainly
static and cannot be easilly configured to meet the needs of other uses cases
besides the benchmarks they have been designed for. In this regard, the approaches
like those proposed in ~\cite{Nettleton2016} and ~\cite{Sukthankar-SocialInfo2014} offer a
more flexible and understandable configuration process to tweak the types,
values and correlations between different features.

Regarding the correlation between the underlying graph structure and the node
features, approaches such as LDBC-SNB, S3G2 or ~\cite{Nettleton2016}
take into account this aspect and the generated graphs have ralistic
structural properties while similar nodes have a larger probability of being
connected. However, their approach seems to be more based on
intuition and common sense than to be backed up by an any study of how the
relation between structure and attributes showcase in real social networks. In
this reagard, this remains as a clear open challenge for social network
generators.

Finally, scalability is another aspect to be considered in social network graph
generators. LDBC-SNB and S3G2 are engineered with this in mind, thus they
provide a way to scale to billions of nodes and edges. This is not the case for
the other generators, which can make them impractical if our goal is to generate
real sized social network graphs.

>>>>

On-line social networks, like Facebook, Twitter, or LinkedIn, have become a
phenomenon used by billions of people every day and thus providing extremely
useful information for various domains. However, an analysis of such type of
graphs has to cope with two problems: (1) availability of the data and (2)
privacy of the data. Hence, data generators which provide realistic synthetic
social network graphs are in a great demand.

In general, any analysis of social networks identifies their various specific
features~\cite{Chakrabarti:2006:GML:1132952.1132954}. For example, a social
graph often has high \emph{clustering coefficient}, i.e., the degree of
transitivity of a graph. Or, its diameter, i.e., the longest shortest path
amongst some fraction (e.g. 90\%) of all connected nodes, is usually low due to
weak ties joining faraway cliques.

Another important aspect of social networks is the community effect. A detailed
study of structure of communities in 70 real-world networks is provided, e.g.,
in~\cite{Leskovec:2008:SPC:1367497.1367591}.
~\cite{Prat-Perez:2014:CSS:2621934.2621942} analyzed the structure of
communities (clustering coefficient, triangle participation ratio, bridges,
diameter, conductance and size) in both real-world graphs and outputs of
existing graph generators such as LFR~\cite{PhysRevE.78.046110} and the
LDBC SNB~\cite{Erling:2015:LSN:2723372.2742786}. They found out that discovered
communities  in different graphs have common distributions and that communities
in a single graph have diverse nature and are difficult to fit with a single
model.

@@

\paragraph{Realistic Social Network}
~\cite{Barrett:2009:GAL:1995456.1995598} focused on the construction of
realistic social networks. For this purpose the authors combine both public and
private data sets with large-scale agent-based techniques. The process works as
follows: In the first step it generates  synthetic data by combining public and
commercial databases. In the second step, it determines a set of activity
templates. A 24-hour activity sequence including geolocations is assigned to
each synthetic individual. To demonstrate the approach, the authors create a
synthetic US population consisting of people and households together with
respective geolocations. For this purpose the authors combine simulation and
data fusion techniques utilizing various real-world data sources such as U.S.
Census data, responses to an activity survey or a time-use survey.
%Demographic information for each person and location, a minute-by-minute
%schedule of each person's activities, and the locations where these activities
%take place is generated by a combination of simulation and data fusion
%techniques.
The result is captured by a dynamic social contact network. Similar methods for
agent-based strategies have been reported
in~\cite{Bernstein:2013:SAS:2499604.2499609}.

@@

\paragraph{Cloning of Social Networks} ~\cite{Sukthankar-SocialInfo2014}
introduces two synthetic generators to reproduce two characteristics typically
observed in social networks: node features and multiple link types. Both
generators extend  the generator proposed by~\cite{wang2011leveraging}.
which starts with a small number of
nodes and new nodes are added until the network reaches the required number. It
has two basic parameters: homophily and link density. A high \emph{homophily}
value reflects that links have higher chances to be established among the nodes with the
same label; these labels can be considered as being equivalent to community
membership.

The first proposed generator is Attribute Synthetic Generator (ASG), used for
reproducing the node feature distribution of standard networks and rewiring the
network to preferentially connect nodes that exhibit a high feature similarity.
The network is initialized with a group of three nodes. New nodes and links
are added to the network based on link density, homophily, and feature
similarity. As new nodes are created, their labels are assigned based on the
prior label distribution. After the network has reached the same number of nodes
as the original social media dataset, each node initially receives a random
attribute assignment. Then a stochastic optimization process is used to move the
initial assignments closer to the target distribution extracted from social
media dataset using the Particle Swarm Optimization algorithm. The tuned
attributes are then used to add additional links to the network based on the
feature similarity parameter -- a source node is selected randomly and connected
to the most similar node. The second proposed generator, so-called Multi-Link Generator
(MLG), further  uses link co-occurrence statistics from the original dataset to
create a multiplex network. MLG uses the same network growth process as ASG.
Based on the link density parameter, either a new node is generated with a label
based on the label distribution of the target dataset or a new link is created
between two existing nodes.

@@

\paragraph{Towards More Realistic Data} \cite{Nettleton2016} argued that the majority of existing works focuses on
topology generation which approximates the characteristics of a real-world social
network (such as a small graph diameter, skew degree
distribution, small average path length, and community structures); however,  this is usually done without any data. Hence, they
introduced a general stochastic modeling approach that enables the users to
populate a graph topology with data. The approach has three steps: (1) topology
generation (using R-MAT) plus community identification using the Louvain
method~\cite{1742-5468-2008-10-P10008} or usage of a real-world topology from
SNAP\footnote{\url{https://snap.stanford.edu/data/}}, (2) data definition
that describes distribution profiles, attribute value definitions, using a
parameterizable set of data propagation rules and affinities, and (3) data
population.

\paragraph{Strengths and Weaknesses of Social Network Generators}
Compared to more general graph generators, social network generators focus
mainly on reproducing intra- and inter-node feature correlations.
Among existing generators, LDBC SNB and S3G2 look
like the most advanced ones in terms of the complexity of the generated graph
and the amount of features and correlations they can generate, while providing a
large degree of scalability. Their generation process is based on input dictionaries and have configuration files that
allow tweaking  many parameters of the generated graphs, but their schema is mainly
static and cannot be easily configured to meet the needs of other uses cases
besides the benchmarks they have been designed for. In this regard, the approaches
like those proposed in~\cite{Nettleton2016} and~\cite{Sukthankar-SocialInfo2014} offer a
more flexible and understandable configuration process to tweak the types,
values, and correlations between different features.

Regarding the correlation between the underlying graph structure and the node
features, approaches such as LDBC SNB, S3G2 or ~\cite{Nettleton2016}
take into account this aspect and the generated graphs have realistic
structural properties while similar nodes have a larger probability of being
connected. However, their approach seems to be more based on
intuition and common sense than to be backed up by  any study of how the
relation between structure and attributes showcase in real social networks. In
this regard, this remains as a clear open challenge for social network
generators.

Finally, scalability is another aspect to be considered in social network graph
generators. LDBC SNB and S3G2 are engineered with this in mind, thus they
provide a way to scale to billions of nodes and edges. This is not the case for
the other generators, which can make them impractical if our goal is to generate
real sized social network graphs.

### 2019-07-23 generators_community_detection.tex
Consolidation of Sections 3-5

Community detection is one of the many graph analytics algorithms typically used
in domains such as social networks or bioinformatics. \emph{Communities} are groups of nodes that are highly connected among them, while being scarcely connected to
the rest of the graph. Such communities emerge from the fact that real graphs
are not random, but follow real-world dynamics that make similar entities to
have a larger probability to be connected. As a consequence, detected
communities are used to reveal vertices with similar characteristics, for
instance to discover functionally equivalent proteins in protein-protein
interaction networks, or persons with similar interests in social networks. Such
applications have made community detection a hot topic during the last fifteen
years with tens of developed algorithms and detection
strategies~\cite{doi:10.1002/wics.1403,Kim:2015:CDM:2854006.2854013}. For
comparing the quality of the different proposed techniques, one needs graphs
with \emph{reference communities}, that is, communities known beforehand. Since
it is very difficult to have large real graphs with reference communities
(mainly because these would require a manual labeling), graphs for benchmarking
community detection algorithms are typically generated synthetically.

\paragraph{Danon et al.} The first attempts to compare community detection algorithms using synthetic
graphs proposed the use of random graphs composed by several Erd\"{o}s-R\'{e}nyi
subgraphs, connected more internally than externally~\cite{danon2005comparing}.
Each of these subgraphs has the same size and the same internal/external density
of edges. However, such graphs miss the realism observed in real graphs, where
communities are of different sizes and densities, thus several proposals exist
to overcome such an issue.

@@

\paragraph{Strengths and Weaknesses of Community Detection Generators}
Besides synthetic graph generators, Yang and Leskovec~\cite{yang2015defining}
proposed the use of real-world graphs with explicit group annotations (e.g.,
forums in a social network, categories of products, etc.) to infer what they
call ``meta-communities'', and use them to evaluate overlapping community
detection algorithms. However, a recent study from Hric, Darst and
Fortunato~\cite{hric2014community} reveal a loose correspondence between
communities (the authors refer to them as structural communities) and
meta-communities.  This result reveals that  algorithms working for structural
communities do not work well for finding meta-communities and vice versa,
suggesting significantly different underlying characteristics
between the two types of communities, which are yet to be
identified.

In this regard and to the best of our knowledge, there does not exist a set of
generators to generate graphs with meta-communities for community detection
algorithm benchmarking. The closest one is the LDBC-SNB data
generator~\cite{Erling:2015:LSN:2723372.2742786} which has been provided by the
generation of groups of users in the social network. Even though the generation
process does not specifically enforce the generation of groups
(meta-communities) for benchmarking community detection algorithms, the study
from Prat-Pe\'rez and Dominguez-Sal reveals that these groups are more similar
to the real meta-communities than those structural communities generated by the
LFR benchmark~\cite{Prat-Perez:2014:CSS:2621934.2621942}.

>>>>

Community detection is one of the many graph analytics algorithms typically used
in domains such as social networks or bioinformatics. \emph{Communities} are groups of nodes that are highly mutually connected, while being scarcely connected to
the rest of the graph. Such communities emerge from the fact that real-world graphs
are not random, but follow real-world dynamics that make similar entities to
have a larger probability to be connected. As a consequence, detected
communities are used to reveal vertices with similar characteristics, for
instance to discover functionally equivalent proteins in protein-to-protein
interaction networks, or persons with similar interests in social networks. Such
applications have made community detection a hot topic during the last 15
years with tens of developed algorithms and detection
strategies~\cite{doi:10.1002/wics.1403,Kim:2015:CDM:2854006.2854013}. For
comparing the quality of the different proposed techniques, one needs graphs
with \emph{reference communities}, that is, communities known beforehand. Since
it is very difficult to have large real-world graphs with reference communities
(mainly because these would require a manual labeling), graphs for benchmarking
community detection algorithms are typically generated synthetically.

\paragraph{Danon et al.} The first attempts to compare community detection algorithms using synthetic
graphs proposed the use of random graphs composed by several Erd\"{o}s-R\'{e}nyi
subgraphs, connected more internally than externally~\cite{danon2005comparing}.
Each of these subgraphs has the same size and the same internal/external density
of edges. However, such graphs miss the realism observed in real-world graphs, where
communities are of different sizes and densities, thus several proposals exist
to overcome such an issue.

@@

\paragraph{Stochastic Block Models} Another popular family of generators
widely used in the community detection field are the stochastic block
models~\cite{holland1983stochastic}. In such models, the community structure of
the graph is typically defined as an array of $n$ community or cluster sizes
and a density square matrix of size $n \times n$ containing the density of
intra-cluster edges (in the diagonal of the matrix) and the density of
inter-cluster edges. Then, a stochastic procedure is run to sample graphs
from such array and matrix, using the sizes to compute the possible edges
and the densities as probabilities of such edges to exist. The popularity of
these methods stem from its simplicity and scalability, which makes them
suitable for generating large graphs fast and in a distributed environments,
provided that the density matrix is sparse (as it happens in most of real-world
                                            world graphs). Moreover, given
that the generation process of such models is mathematically tractable, they
are typically used to analyze the limitations of community detection
algorithms such as those based on modularity
optimization~\cite{fortunato2007resolution} or based on
triads~\cite{prat2016put}. Extensions of such models exist, such as the
Mixed Membership Stochastic Block Model~\cite{airoldi2008mixed}, for
overlapping communities.

\paragraph{Strengths and Weaknesses of Community Detection Generators}
Besides synthetic graph generators, Yang and Leskovec~\cite{yang2015defining}
proposed the use of real-world graphs with explicit group annotations (e.g.,
forums in a social network, categories of products, etc.) to infer what they
call \emph{meta-communities}, and use them to evaluate overlapping community
detection algorithms. However, a recent study from Hric, Darst and
Fortunato~\cite{hric2014community} reveal a loose correspondence between
communities (the authors refer to them as \emph{structural communities}) and
meta-communities.  This result reveals that  algorithms working for structural
communities do not work well for finding meta-communities and vice versa,
suggesting significantly different underlying characteristics
between the two types of communities, which are yet to be
identified.

In this regard and to the best of our knowledge, there does not exist a set of
generators to generate graphs with meta-communities for community detection
algorithm benchmarking. The closest one is the LDBC SNB data
generator which has been provided by the
generation of groups of users in the social network. Even though the generation
process does not specifically enforce the generation of groups
(meta-communities) for benchmarking community detection algorithms, the study~\cite{Prat-Perez:2014:CSS:2621934.2621942} reveals that these groups are more similar
to the real meta-communities than those structural communities generated by the
LFR benchmark.

### 2019-07-23 generators_general.tex
Consolidation of Sections 3-5

\paragraph{Strengths and Weaknesses of General Graph Generators}
In general, existing general graph generators produce graphs with the following
properties: skewed degree distribution (e.g. powerlay), small diameter, large
clustering coefficient, a large largest connected component and some degree of
community structure. Degree distribution can be typically configured while other
properties are just a result of the genration process and cannot be controlled
by any means.

This is not the case of BTER and Darwini which besides the degree distribution,
they also allow tunning the clustering coefficient. However, some use cases
demand for more control on the characteristics of the generated graphs.
This is the case, for instance, of benchmarking, where the underlying graph
structure has direct implications to the performance of the graph algorithms
to run. For this reason, the design of graph generators with the capability of fine
tunning the characteristics of the generated graphs still remains as an open
challenge. The question that remains though is what characteristics to tune,
and which are the algorithms whose depends on such characteristics.

>>>>

\paragraph{Strengths and Weaknesses of General Graph Generators}
In general, existing general graph generators produce graphs with the following
properties: skewed degree distribution (e.g., powerlay), small diameter, large
clustering coefficient, a large largest connected component, and some degree of
community structure. Degree distribution can be typically configured while other
properties are just a result of the generation process and cannot be controlled
by any means.

This is not the case of BTER and Darwini, which besides the degree distribution,
they also allow tuning of the clustering coefficient. However, some use cases
demand for more control on the characteristics of the generated graphs.
This is the case, for instance, of benchmarking, where the underlying graph
structure has direct implications to the performance of the graph algorithms
to run. For this reason, the design of graph generators with the capability of fine
tuning the characteristics of the generated graphs still remains as an open
challenge. The questions that remain are what characteristics to tune
and which are the algorithms that depend on such characteristics.
